VIPRE: Virtualized Intermittently-Powered Runtime Environment

### 2017-07-11 paper/background.tex
Expanding background section (part 1 - harvesting - kind of ready)

Many platforms enable intermittent computation already. Those include computation RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}. Other intermittently-powered platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}. In all of the above, the main source of energy harvested is the electromagnetic radiation in the radio frequency range (ambient transmitters such as high power TV transmitters~\cite{liu_sigcomm_2013} or dedicated RFID antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}). Naturally, other forms of energy harvesting sources exist, including temperature gradient, light/sun radiation, vibrations, (micro-)motions, and body fluid flow (blood, gastric acid). The reader is referred to example extensive surveys discussing energy harvesting for low-power embedded systems~\cite{paradiso_pvc_2005,soyata_csm_2016,prasad_comst_2014,ku_cst_2016}.

Application include long-term battery-less monitoring. One example is the distributed monitoring of moisture of large plant fields during vegetation~\cite{}. Other applications include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, insect monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015}  and digestible~\cite{nadeau_naturebio_2017} sensors.

\begin{table}
	\begin{tabular}{|c|c|}
	\hline
	Platform Name & Storage capacitor size \\
	\hline \hline
	NeuralWISP~\cite{} & 100\,$\mu$F \\
	BioImpedance~\cite{} & 20\,$\mu$F \\
	WISPCam~\cite{} & 6.08\,mF \\ %tested [11.24, 17.45, 21.98]\,mF
	NFC-WISP~\cite{zhao_rfid_2015} & 300\,$\mu$F \\
	WISP~\cite{} & 47\,$\mu$F \\
	\hline
	\end{tabular} 
\caption{Comparison of storage sizes for various battery-less platforms.}
\label{table:capacitor}
\end{table}

%DragonFly~\cite{} Not reported
%Ambient Backscatter~\cite{,} Not reported
%Fasense~\cite{} Not reported
%Battery-less phone~\cite{} Not reported

Energy provided from the ambient is not stable and difficult to predict accurately. Combined with the fact that energy supply is small, there is little leeway to store enough of energy to guarantee prolonged periods of computation. Moreover, each battery-less platform has different storage sizes, refer to Table~\ref{table:capacitor}. Energy breaks happen every hundred of milliseconds~\cite{soyata_csm_2016}.

Solution to the problem is to divide the program into smaller parts that guarantee execution within a discharge region and track the state of the program in-between the program part execution times.

@@

Various capacitor sizes of intermittent device result in a program being executed

@@

>>>>

Given current technology development, battery-less systems are best suited for very long-term sensing and monitoring where access to recharge is either prohibitive or impossible. These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, insect monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015}  and digestible~\cite{nadeau_naturebio_2017} sensors.

Many platforms enable intermittent, battery-less computation. For instance, computation RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}. Other intermittently-powered platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}. In all of the above, the main source of energy harvested is the electromagnetic radiation in the radio frequency range (ambient transmitters such as high power TV transmitters~\cite{liu_sigcomm_2013} or dedicated RFID antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}). Naturally, other forms of energy harvesting sources exist, including temperature gradient, (micro-)motions, light/sun radiation, vibrations, and body fluid flow (blood, gastric acid). The reader is referred to example extensive surveys discussing energy harvesting for low-power embedded systems~\cite{paradiso_pvc_2005,soyata_csm_2016,prasad_comst_2014,ku_cst_2016}.

\begin{table}
	\begin{tabular}{|c|c|}
	\hline
	Platform name & Storage capacitor size \\
	\hline \hline
	Moo~\cite{moo} & 0.1\,F \\
	WISPCam~\cite{naderiparizi_rfid_2015} & 6.08\,mF \\ %tested [11.24, 17.45, 21.98]\,mF
	NFC-WISP~\cite{zhao_rfid_2015} & 300\,$\mu$F \\
	NeuralWISP~\cite{holleman_biocas_2008} & 100\,$\mu$F \\
	WISP~\cite{wisp5} & 47\,$\mu$F \\
	{\em BioImpedance} sensor~\cite{rodriguez_tbcs_2015} & 20\,$\mu$F \\
	{\em Ingestible} sensor~\cite{nadeau_naturebio_2017} & 220\,$\mu$F\\
	\hline
	\end{tabular} 
\caption{Comparison of {\em default} energy storage sizes for various battery-less platforms. Observe a huge variation in storage capacity. We note that values for other representative platforms~\cite{medusa_farsens_2017,talla_imwut_2017,liu_sigcomm_2013,parks_sigcomm_2014} were not reported in their respective papers.}
\label{table:capacitor}
\end{table}

The biggest technical problem with energy provided from the ambient to battery-less platforms is that harvested energy is not stable and difficult to predict accurately. Combined with the fact that energy supply is small, there is little leeway to store enough of energy to guarantee prolonged periods of computation. Energy breaks happen every hundred of milliseconds, refer again to e.g.~\cite[Fig. 1]{mementos}. This causes each program running on such platform to restart. Solution to the problem is to divide the program into smaller parts that guarantee execution within a discharge region and track the state of the program in-between execution times.

@@

Moreover, each battery-less platform has different storage sizes, refer to Table~\ref{table:capacitor}, which causes different charge/discharge periods. This result in different execution times. In the worst case the size of the capacitor might be too small to execute one task written by the programmer.

@@

%prior systems do fine grained task division (virtualization) {clank,aplaca}
